Supporting Information

### 2018-04-09 SupportingInformation.Rmd
a few developments in the Taylor's law stuff

The model-averaged (averaging performed using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistics are in Tables \ref{tab_pns_LTdep_AICw} and \ref{tab_pns_UTdep_AICw}, respectively. These values were exclusively positive, thereby
differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data - these data
have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis.

@@

Mean values of our statistics, as well as mean Kendall and Spearman
correlations, were computed across all pairs of sampling locations, and confidence intervals of these means were computed via a
spatial resampling scheme, as described in Methods (Table \ref{tab_pns_resamp}).
Confidence intervals of $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$
always excluded $0$ and mean values were positive, indicating tail dependence.
Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$
also all were strictly positive, indicating asymmetry of tail dependence, a feature that data with normal copula
structure do not have, providing further evidence answering the third part of question 1.

@@

Plotting Spearman or Kendall correlations between time series for pairs of locations as a function of great circle distance
between the locations (Figure \ref{fig_pns} A, B) showed the typical decline of correlation with increasing
distance. Plotting $\cor_l$ , $\cor_u$ , $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear to decline with distance (Figure \ref{fig_pns} C-F). $\Dsq_u$ and $\Dsq_l$ appear to increase with distance, as expected given
their opposite meaning (Figure \ref{fig_pns} G-H). Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Figure \ref{fig_pns} I-K) did not show a decline with distance, but did demonstrate visually the 
statistical result of Table \ref{tab_pns_resamp} that averages of these 
quantities across all pairs of sampling locations were significantly greater 
than zero, showing asymmetry of tail dependence.

@@

# Question 3 analyses for each dataset
Recall that question 3 is: What are the consequences of non-Gaussian copula structure and tail dependence for our understanding of biology and for applications?

@@

#***generate the starting data, which has normal copula structure
#because it comes from a multivariate normal
sig<-matrix(rho,n,n)
diag(sig)<-1
m<-rmvnorm(N*numdat,mean=rep(0,n),sigma=sig)
m<-array(m,c(N,numdat,n))
m<-aperm(m,c(1,3,2))
gshape<-7.5
gscale<-1 #gamma parameters
cop_to_marg_trans<-function(x)
{
  return(qgamma(x,shape=gshape,scale=gscale)) #this is where we choose the marginals,
                                  #should be something skewed
}
marg_to_cop_trans<-function(x) #transforms to copula space
{
  return(pgamma(x,shape=gshape,scale=gscale))
}
m<-cop_to_marg_trans(pnorm(m))

@@

<!--do a TL analysis on the data just created-->
```{r tl_analysis, echo=F, cache=T, cache.extra=list(mtime("TaylorsLawTools.R"),seed,mtime("copsurrognd.R"))}
source("copsurrognd.R")
source("TaylorsLawTools.R")
# 
# #We are only doing spatial taylor's law because we know a priori temporal
# #won't give anything interesting for the surrogates we use
# 
 kend.stats.res<-list()
 kend.points.res<-list()
 spea.stats.res<-list()
 spea.points.res<-list()
 for (counter in 1:length(kend.surrog))
 {
   ks<-kend.surrog[[counter]]
   ss<-spea.surrog[[counter]]
   kend.stats.res[[counter]]<-matrix(NA,5,numdat)
   spea.stats.res[[counter]]<-matrix(NA,5,numdat)
   kend.points.res[[counter]]<-array(NA,c(N,2,numdat))
   spea.points.res[[counter]]<-array(NA,c(N,2,numdat))
   for (sc in 1:numdat)
   {
     khold<-tl_stats(ks[,,sc])
     kend.stats.res[[counter]][,sc]<-khold$s.stats #spatial TL only
     rownames(kend.stats.res[[counter]])<-names(khold$s.stats)
     kend.points.res[[counter]][,,sc]<-as.matrix(khold$s.points)
# 
     shold<-tl_stats(ss[,,sc])
     spea.stats.res[[counter]][,sc]<-shold$s.stats #spatial TL only
     rownames(spea.stats.res[[counter]])<-names(shold$s.stats)
     spea.points.res[[counter]][,,sc]<-as.matrix(shold$s.points)
   }
 }
 names(kend.stats.res)<-names(kend.surrog)
 names(kend.points.res)<-names(kend.surrog)
 names(spea.stats.res)<-names(spea.surrog)
 names(spea.points.res)<-names(spea.surrog)
```

<!--Now make some plots-->
```{r taylor_plots, echo=T}
##plot TL for clayton
xlimits<-range(kend.points.res[[1]][,1,],kend.points.res[[2]][,1,],kend.points.res[[3]][,1,],kend.points.res[[4]][,1,])
ylimits<-range(kend.points.res[[1]][,2,],kend.points.res[[2]][,2,],kend.points.res[[3]][,2,],kend.points.res[[4]][,2,])
par(mfrow=c(1,1))
plot(kend.points.res[[1]][,1,],kend.points.res[[1]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits)  
```

```{r, echo=T}
##plot TL for normal copula
plot(kend.points.res[[2]][,1,],kend.points.res[[2]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits)  
```

```{r, echo=T}
#plot TL for frank
plot(kend.points.res[[3]][,1,],kend.points.res[[3]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits)  
```

```{r, echo=T}
#plot TL for survival clayton
plot(kend.points.res[[4]][,1,],kend.points.res[[4]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits)  
```

>>>>

The model-averaged (averaging performed using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistics are in Tables \ref{tab_pns_LTdep_AICw} and\ref{tab_pns_UTdep_AICw}, respectively. These values were exclusively positive, thereby
differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data - these data
have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis.

@@

Mean values of our statistics, as well as mean Kendall and Spearman
correlations, were computed across all pairs of sampling locations, and confidence intervals of these means were computed via a
spatial resampling scheme, as described in Methods (Table \ref{tab_pns_resamp}).
Confidence intervals of $\cor_l$","$\cor_u$","$\Ps_l$","$\Ps_u$","$\Dsq_l$","$\Dsq_u$"
always excluded $0$ and mean values were positive, indicating tail dependence.
Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$
also all were strictly positive, indicating asymmetry of tail dependence, a feature that data with normal copula
structure do not have, providing further evidence answering the third part of question 1.

@@

Plotting Spearman or Kendall correlations between time series for pairs of locations as a function of great circle distance
between the locations (Figure \ref{fig_pns} A, B) showed the typical decline of correlation with increasing
distance. Plotting $\cor_l$ , $\cor_u$ , $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear to decline with distance (Figure \ref{fig_pns} C-F). $\Dsq_u$ and $\Dsq_l$ appear to increase with distance, as expected given
their opposite meaning (Figure \ref{fig_pns} G-H). Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Figure \ref{fig_pns} I-K) did not show a decline with distance, but did demonstrate visuaally the 
statistical result of Table \ref{tab_pns_resamp} that averages of these 
quantities across all pairs of sampling locations were significantly greater 
than zero, showing asymmetry of tail dependence.

@@

# Question 3 analyses for each dataset and theoretical analyses
Recall that question 3 is: What are the consequences of non-Gaussian copula structure and tail dependence for our understanding of biology and for applications?

@@

#***generate the starting data, which has normal copula structure
#because it comes originally from a multivariate normal, before 
#transformation
sig<-matrix(rho,n,n)
diag(sig)<-1
m<-rmvnorm(N*numdat,mean=rep(0,n),sigma=sig)
m<-array(m,c(N,numdat,n))
m<-aperm(m,c(1,3,2))
gshape<-7.5
gscale<-1 #gamma parameters
cop_to_marg_trans<-function(x)
{
  return(qgamma(x,shape=gshape,scale=gscale)) #this is where we choose the marginals,
                                  #should be something skewed
}
marg_to_cop_trans<-function(x) #transforms to copula space
{
  return(pgamma(x,shape=gshape,scale=gscale))
}
m<-cop_to_marg_trans(pnorm(m))

@@

<!--do a TL analysis on the data just created-->
```{r tl_analysis, echo=F, cache=T, cache.extra=list(kend.surrog,spea.surrog,mtime("TaylorsLawTools.R"),seed,mtime("copsurrognd.R"))}
source("copsurrognd.R")
source("TaylorsLawTools.R")

#We are only doing spatial taylor's law because we know a priori temporal
#won't give anything interesting for the surrogates we use

kend.stats.res<-list()
kend.points.res<-list()
spea.stats.res<-list()
spea.points.res<-list()
for (counter in 1:length(kend.surrog))
{
 ks<-kend.surrog[[counter]]
 ss<-spea.surrog[[counter]]
 kend.stats.res[[counter]]<-matrix(NA,5,numdat)
 spea.stats.res[[counter]]<-matrix(NA,5,numdat)
 kend.points.res[[counter]]<-array(NA,c(N,2,numdat))
 spea.points.res[[counter]]<-array(NA,c(N,2,numdat))
 for (sc in 1:numdat)
 {
   khold<-tl_stats(ks[,,sc])
   kend.stats.res[[counter]][,sc]<-khold$s.stats #spatial TL only
   rownames(kend.stats.res[[counter]])<-names(khold$s.stats)
   kend.points.res[[counter]][,,sc]<-as.matrix(khold$s.points)

   shold<-tl_stats(ss[,,sc])
   spea.stats.res[[counter]][,sc]<-shold$s.stats #spatial TL only
   rownames(spea.stats.res[[counter]])<-names(shold$s.stats)
   spea.points.res[[counter]][,,sc]<-as.matrix(shold$s.points)
 }
}
names(kend.stats.res)<-names(kend.surrog)
names(kend.points.res)<-names(kend.surrog)
names(spea.stats.res)<-names(spea.surrog)
names(spea.points.res)<-names(spea.surrog)
```

<!--Now make some plots, these one for kendall surrogates-->
```{r taylor_plots, echo=F}
##plot TL for clayton
xlimits<-range(kend.points.res[[1]][,1,],kend.points.res[[2]][,1,],kend.points.res[[3]][,1,],kend.points.res[[4]][,1,])
ylimits<-range(kend.points.res[[1]][,2,],kend.points.res[[2]][,2,],kend.points.res[[3]][,2,],kend.points.res[[4]][,2,])
par(mfrow=c(1,1),mar=c(4,4,1,1) + 0.1)
plot(kend.points.res[[1]][,1,],kend.points.res[[1]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")))  
points(kend.points.res[[1]][,1,1],kend.points.res[[1]][,2,1],type='p',pch=20,cex=.75,col='red')
points(kend.points.res[[1]][,1,2],kend.points.res[[1]][,2,2],type='p',pch=20,cex=.75,col='green')

##plot TL for normal copula
plot(kend.points.res[[2]][,1,],kend.points.res[[2]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")))  
points(kend.points.res[[2]][,1,1],kend.points.res[[2]][,2,1],type='p',pch=20,cex=.75,col='red')
points(kend.points.res[[2]][,1,2],kend.points.res[[2]][,2,2],type='p',pch=20,cex=.75,col='green')

#plot TL for frank
plot(kend.points.res[[3]][,1,],kend.points.res[[3]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")))  
points(kend.points.res[[3]][,1,1],kend.points.res[[3]][,2,1],type='p',pch=20,cex=.75,col='red')
points(kend.points.res[[3]][,1,2],kend.points.res[[3]][,2,2],type='p',pch=20,cex=.75,col='green')

#plot TL for survival clayton
plot(kend.points.res[[4]][,1,],kend.points.res[[4]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")))  
points(kend.points.res[[4]][,1,1],kend.points.res[[4]][,2,1],type='p',pch=20,cex=.75,col='red')
points(kend.points.res[[4]][,1,2],kend.points.res[[4]][,2,2],type='p',pch=20,cex=.75,col='green')

#prepare for plots coming below
s.lin<-cbind(kend.stats.res[[1]][1,],kend.stats.res[[2]][1,],kend.stats.res[[3]][1,],kend.stats.res[[4]][1,])
s.hom<-cbind(kend.stats.res[[1]][2,],kend.stats.res[[2]][2,],kend.stats.res[[3]][2,],kend.stats.res[[4]][2,])
s.rmse<-cbind(kend.stats.res[[1]][3,],kend.stats.res[[2]][3,],kend.stats.res[[3]][3,],kend.stats.res[[4]][3,])
s.int<-cbind(kend.stats.res[[1]][4,],kend.stats.res[[2]][4,],kend.stats.res[[3]][4,],kend.stats.res[[4]][4,])
s.slope<-cbind(kend.stats.res[[1]][5,],kend.stats.res[[2]][5,],kend.stats.res[[3]][5,],kend.stats.res[[4]][5,])

#plot s.lin distr for each set of surrogates
#***DAN: x axis needs to be redone to show the copula names used in the surrogates
par(mfrow=c(1,1),mar=c(4,4,1,1) + 0.1)
x<-1:4
plot(x,s.lin[1,],type="b",ylim=range(s.lin),xlab="Set of surrogates",ylab="Linearity test p-vals")
for (counter in 2:dim(s.lin)[1])
{
  lines(x,s.lin[counter,],type='b')
}

#plot s.hom distr for each set of surrogates
#***DAN: x axis needs to be redone to show the copula names used in the surrogates
x<-1:4
plot(x,s.hom[1,],type="b",ylim=range(s.hom),xlab="Set of surrogates",ylab="Homoskedasticity test p-vals")
for (counter in 2:dim(s.slope)[1])
{
  lines(x,s.hom[counter,],type='b')
}

#plot s.rmse distr for each set of surrogates
#***DAN: x axis needs to be redone to show the copula names used in the surrogates
x<-1:4
plot(x,s.rmse[1,],type="b",ylim=range(s.rmse),xlab="Set of surrogates",ylab="Spatial TL rmse")
for (counter in 2:dim(s.rmse)[1])
{
  lines(x,s.rmse[counter,],type='b')
}

#plot s.int distr for each set of surrogates
#***DAN: x axis needs to be redone to show the copula names used in the surrogates
x<-1:4
plot(x,s.int[1,],type="b",ylim=range(s.int),xlab="Set of surrogates",ylab="Spatial TL intercept")
for (counter in 2:dim(s.int)[1])
{
  lines(x,s.int[counter,],type='b')
}

#plot s.slope distr for each set of surrogates
#***DAN: x axis needs to be redone to show the copula names used in the surrogates
x<-1:4
plot(x,s.slope[1,],type="b",ylim=range(s.slope),xlab="Set of surrogates",ylab="Spatial TL slope")
for (counter in 2:dim(s.slope)[1])
{
  lines(x,s.slope[counter,],type='b')
}
```
